Comparing R packages for calculation of persistent homology

### 2020-03-08 tdabench-paper/tdabench.Rmd
revised mem1 text

\subsubsection{Memory Benchmarks}
As point cloud number increases, Rips complexes show a more dramatic rise in run time than Alpha complexes. 
This is predicted by Rips theoretical exponential complexity and Alpha complex's theoretical polynomial complexity. 
In both complexes, the boundary matrix is used to calculate persistent homology (explained in detail in the Mathematics section). 
The library GUDHI provides functions to return the boundary matrix, which allows us to measure object size.

Rips complexes show a rapid growth in object size as the point cloud number increases in figure \ref{fig:mem1}. 
The growth pattern differences mirror our findings for run time growth with increasing point cloud size. 
Alpha complexes show a linear growth in memory complexity whereas Rips complexes show a polynomial growth. 
However, calculating the object size reveals some interesting differences in what parameters determine the object size of Rips vs Alpha complexes.

@@

\begin{figure}
  \centering
  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig5a.png}
  \caption{Persistent Homology Boundary Matrix for Rips vs Alpha complex. The x-axis represents the number of points in the point cloud, and y-axis represents object size. The type of complex is indicated by color. Maximum feature dimensions were calculated. Panel 1 = Comparing object size on 3D annulus. Panel 2 = Comparing object size on sphere. Panel 3 = Comparing object size on torus. Panel 4 = Comparing object size on 3D box. Similar to runtime growth, Alpha complex showed a linear pattern and Rips complex showed a log-log linear pattern. Regression results can be found in supplemental data on the paper Github}
\label{fig:mem1}
\end{figure}

@@

<!--Include somewhere that differences are b/n implementations (practical) not necessarily mathematical structures (theory), so results can change as implementations change (whether b/n libraries or b/n complices)-->

>>>>

Vietoris-Rips and alpha complices differ not only in runtime, but also in memory usage.
GUDHI conveniently allows measurement of the simplicial complex boundary matrix, a reasonable proxy for total memory usage in calculation of persistent homology.
Since Ripser and Dionysus do not have similar capabilities, they are excluded from memory benchmarks.
Vietoris-Rips complices show a rapid growth in object size as point cloud size increases; in contrast, alpha complices consistently maintain low, linearly-growing memory use (Figure \ref{fig:mem1}).
The growth pattern differences mirror findings for runtime growth (Figure \ref{fig:ann}).
Interestingly, different factors determine memory usage in the two types of simplicial complex.

<!--BELONGS IN DISCUSSION
This is predicted by Vietoris-Rips theoretical exponential complexity and Alpha complex's theoretical polynomial complexity.-->

@@

\begin{figure}
  \centering
  \includegraphics[height=2.5in]{../Figures/Final_Figures/fig5a.png}
  \caption{
    \textbf{Comparing memory usage of persistent homology calculation using a Vietoris-Rips versus an alpha complex.}
    Mean boundary matrix object size, a proxy for memory usage, plotted against point cloud size varying by point cloud structure (facet) and simplicial complex type (color, shape).
    Data dimension and feature dimension were held constant at 3 and 2, respectively.
    Point cloud structures for which memory usage was calculated were 3-dimensional annulus (top-left), 3-dimensional sphere (top-right), torus (bottom-left), and 3-dimensional cube (bottom-right).
    Similar to runtime results, alpha complex calculation showed linear growth and Vietoris-Rips complex calculation showed polynomial growth (see Supplement for regression details).}
\label{fig:mem1}
\end{figure}

@@

<!--Include somewhere that differences are b/n implementations (practical) not necessarily mathematical structures (theory), so results can change as implementations change (whether b/n libraries or b/n complices)

Ripser memory not measured here b/c R, but refer to Bauer Ripser memory measurements.
-->
